\begin{thebibliography}{99}

\bibitem{rovelli2008loop} C. Rovelli, ``Loop quantum gravity,'' \emph{Living Reviews in Relativity}, vol. 11, no. 5, 2008. https://doi.org/10.12942/lrr-2008-5.

\bibitem{whittaker1951history} E. T. Whittaker, \emph{A History of the Theories of Aether and Electricity}, 2 vols., New York: Dover, 1951--1953.

\bibitem{jacobson2004einstein} T. Jacobson and D. Mattingly, ``Einstein-Aether Theory,'' \emph{Phys. Rev. D}, vol. 70, p. 024003, 2004, arXiv:gr-qc/0007031.

\bibitem{unruh1981experimental} W. G. Unruh, ``Experimental black-hole evaporation?'' \emph{Physical Review Letters}, vol. 46, no. 21, p. 1351, 1981.

\bibitem{steinhauer2016hawking} J. Steinhauer, ``Observation of quantum Hawking radiation and its entanglement in an analogue black hole,'' \emph{Nature Physics}, vol. 12, no. 10, pp. 959--965, 2016.

\bibitem{svancara2024rotating} P. Švančara et al., ``Rotating curved spacetime signatures from a giant quantum vortex,'' \emph{Nature}, vol. 628, no. 8006, pp. 66--70, 2024.

\bibitem{eto2024knots} M. Eto, Y. Hamada, and M. Nitta, ``Tying knots in particle physics,'' arXiv:2407.11731, 2024.

\bibitem{thomson1867vortex} W. Thomson (Lord Kelvin), ``On vortex atoms,'' \emph{Philosophical Magazine}, vol. 34, no. 227, pp. 15--24, 1867.

\bibitem{candelas1985vacuum} P. Candelas, G. T. Horowitz, A. Strominger, and E. Witten, ``Vacuum configurations for superstrings,'' \emph{Nuclear Physics B}, vol. 258, pp. 46--74, 1985.

\bibitem{chamseddine2007gravity} A. H. Chamseddine, A. Connes, and M. Marcolli, ``Gravity and the standard model with neutrino mixing,'' \emph{Advances in Theoretical and Mathematical Physics}, vol. 11, no. 6, pp. 991--1089, 2007.

\bibitem{ashtekar1986new} A. Ashtekar, ``New variables for classical and quantum gravity,'' \emph{Physical Review Letters}, vol. 57, no. 18, p. 2244, 1986.

\bibitem{vortex_dynamics} Saffman, P. G., \emph{Vortex Dynamics}, Cambridge University Press, 1992.

\bibitem{braid_topology} Birman, J. S., \emph{Braids, Links, and Mapping Class Groups}, Princeton University Press, 1974.

\bibitem{quasicrystals} Shechtman, D., et al., ``Metallic Phase with Long-Range Orientational Order and No Translational Symmetry,'' Physical Review Letters, 53, 1951--1953, 1984.

\bibitem{entropy_encoding} MacKay, D. J. C., \emph{Information Theory, Inference, and Learning Algorithms}, Cambridge University Press, 2003.

\end{thebibliography}
