\section*{Forward}

This work emerged from an unusual path that I believe deserves explanation. The core ideas presented here first took shape nearly twenty years ago during my college years, sparked by reading about Nikola Tesla's theoretical writings on the aether. While Tesla's insights were compelling, I lacked the mathematical skills necessary to investigate whether such ideas could be developed into a coherent physical theory.

Recent advances in computational tools and automated reasoning have fundamentally changed what individual researchers can accomplish. These developments enabled me to translate conceptual ideas into rigorous mathematical form—a process that would have been extremely difficult for someone with my background to tackle alone in earlier decades.

I must emphasize that this paper was never supposed to be written. My initial goal was purely to explore these ideas and discover their limits. As a software engineer, I approached this investigation like any complex system—by writing comprehensive tests for every mathematical step. Over the course of this work, I have built hundreds of SymPy verification scripts to ensure dimensional consistency, mathematical accuracy, and proper handling of limiting cases.

What began as personal curiosity gradually revealed unexpected depth. Each apparent dead end, when examined closely, seemed to resolve into deeper consistency. Eventually, the mathematical structure became comprehensive enough that formal documentation became necessary—both to organize my own understanding and to enable others to identify where my reasoning must have gone wrong.

I began this journey expecting to fail and wanting to understand where these ideas would break down. I did not set out to challenge established physics, but rather to explore whether certain concepts could be made mathematically precise. The fact that this exploration has yielded a framework capable of reproducing established gravitational phenomena from fluid-mechanical principles is as surprising to me as it may be to the reader.

Regardless of the ultimate validity of the physical interpretation presented here, I hope that the mathematical connections established may prove valuable to the broader physics community. Even if the aether paradigm is incorrect, these mathematical relationships suggest deeper structural connections that merit investigation.

